Rao-Blackwellized smoother for Conditionally Linear and Gaussian models

### 2016-11-15 TexArxiv/clgm.tex
MAJ

\subsection{Model}
Modeling commodity prices is a crucial step to valuate contingent claims related to energy markets and to optimize storage or extraction strategies. In \cite{gibson:schwartz:1990,schwartz:1997}, the authors proposed a model where the spot price of a commodity$(S_t, t \geq 0)$ depends on a second factor $(\delta_t, t \geq 0)$, referred to as the instantaneous convenience yield. This factor plays the role of dividends in equity markets and models the benefit of holding the physical commodity or the storage and maintenance costs required to keep the commodity. In this model, this convenience yield is described as an Ornstein-Uhlenbeck process:
\rmd S_{t} & = (r-\delta_{t})S_{t}\rmd t+\sigma S_{t} \rmd W_{t}^{1} \eqsp,\\
\rmd \delta_{t} & = \kappa(\alpha-\delta_{t})\rmd t+\eta \rmd W_{t}^{2} \eqsp, \quad \rmd \langle W_t^1, W_t^2\rangle = \rho \rmd t\eqsp,
where the parameter $\left(r, \sigma, \kappa, \alpha, \eta, \rho \right)$ are constant and $( (W_t^1,W_t^2), t \geq 0)$ are standard Brownian motions. This model appear to be too restrictive as energy markets are not likely to revert to a single equilibrium value. This assumption is relaxed using Markov switching models to allow several possible regimes for the spot price and the convenience yield. Following \cite{almansour:2016}, the spot price and convenience yield are described in this paper as:
\rmd S_{t} & = (r-\delta_{t})S_{t}\rmd t+\sigma_{a_t} S_{t}\rmd W_{t}^{1} \eqsp,\\
\rmd\delta_{t} & = \kappa_{a_t}(\alpha_{a_t} -\delta_{t})\rmd t+\eta_{a_t} \rmd W_{t}^{2} \eqsp, \quad \rmd\langle W_t^1, W_t^2\rangle = \rho_{a_t} \rmd t\eqsp,
where $(a_t)_{t\ge 0}$ is a finite state space Markov process.
%We assume that $J=2$, each regime characterizing the slope of the future curve. A backwardation regime (resp. contango regime) is defined as a negative slope (resp. a positive slope).

@@

Z_i = d_{a_{i-1}} + T_{a_{i-1}} Z_{i-1} + H_{a_{i-1}} \varepsilon_i\eqsp,

@@

where (with $\barH_{a_{i-1}} \eqdef H_{a_{i-1}}H'_{a_{i-1}}$ and $\Delta t_i = t_i-t_{i-1}$):

@@

\begin{pmatrix} \left[\mu- \alpha_{a_{i-1}} - \sigma^2_{a_{i-1}}/2 \right] \Delta t_i + \alpha_{a_{i-1}}[1-\rme^{-\kappa_{a_{i-1}} \Delta t_i}]/\kappa_{a_{i-1}}\\
\alpha_{a_{i-1}} [1-\rme^{-\kappa_{a_{i-1}} \Delta t_i}] \end{pmatrix}\eqsp,\\
T_{a_{i-1}} &\eqdef
\begin{pmatrix} 1 & -[1-\rme^{-\kappa_{a_{i-1}} \Delta t_i}]/\kappa_{a_{i-1}} \\ 0 & \rme^{-\kappa_{a_{i-1}} \Delta t_i} \end{pmatrix} \eqsp,\\
\barH_{a_{i-1}}(1,1) &= \sigma^2_{a_{i-1}} \Delta t_i + \eta^2_{a_{i-1}}\left\{\Delta t_i + (1-\rme^{-2\kappa_{a_{i-1}} \Delta t_i})/(2\kappa_{a_{i-1}}) - 2(1-\rme^{-\kappa \Delta t_i})/\kappa_{a_{i-1}}\right\}/\kappa_{a_{i-1}}^2 \\
&\hspace{5.2cm} - 2\rho_{a_{i-1}} \eta_{a_{i-1}} \sigma_{a_{i-1}}\left\{ \Delta t_i - (1-\rme^{-\kappa_{a_{i-1}} \Delta t_i})/\kappa_{a_{i-1}} \right\}/\kappa\eqsp, \\
\barH_{a_{i-1}}(1,2) & = \left(\rho_{a_{i-1}} \eta_{a_{i-1}} \sigma_{a_{i-1}}-\eta^2_{a_{i-1}}/\kappa_{a_{i-1}}\right)\left(1-\rme^{-\kappa_{a_{i-1}}\Delta t_i} \right)/\kappa_{a_{i-1}} \\
&\hspace{5.2cm}+ \eta^2_{a_{i-1}}\left(1-\rme^{-2\kappa_{a_{i-1}} \Delta t_i} \right)/(2\kappa_{a_{i-1}}^2)\eqsp,\\

@@

\barH_{a_{i-1}}(2,2) &= \eta^2_{a_{i-1}}\left( 1-\rme^{-2\kappa_{a_{i-1}}\Delta t_i} \right)/(2\kappa_{a_{i-1}})\eqsp.

@@

Following \cite{almansour:2016}, it is assumed that $\kappa_1 = \kappa_2 = \kappa$ (and therefore $T_1=T_2=T$) to obtain a closed form solution for $F_{t_i,m}$ which may be written:

>>>>

\subsection{Model}
Modeling commodity prices is a crucial step to valuate contingent claims related to energy markets and to optimize storage or extraction strategies. In \cite{gibson:schwartz:1990,schwartz:1997}, the authors proposed a model where the spot price of a commodity$(S_t, t \geq 0)$ depends on a second factor $(\delta_t, t \geq 0)$, referred to as the instantaneous convenience yield. This factor plays the role of dividends in equity markets and models the benefit of holding the physical commodity or the storage and maintenance costs required to keep the commodity. In this model, this convenience yield is described as an Ornstein-Uhlenbeck process:
\begin{align*}
\rmd S_{t} & = (r-\delta_{t})S_{t}\rmd t+\sigma S_{t} \rmd W_{t}^{1} \eqsp,\\
\rmd \delta_{t} & = \kappa(\alpha-\delta_{t})\rmd t+\eta \rmd W_{t}^{2} \eqsp, \quad \rmd \langle W_t^1, W_t^2\rangle = \rho \rmd t\eqsp,
\end{align*}
where the parameter $\left(r, \sigma, \kappa, \alpha, \eta, \rho \right)$ are constant and $( (W_t^1,W_t^2), t \geq 0)$ are standard Brownian motions. This model appear to be too restrictive as energy markets are not likely to revert to a single equilibrium value. This assumption is relaxed using Markov switching models to allow several possible regimes for the spot price and the convenience yield. Following \cite{almansour:2016}, the spot price and convenience yield are described in this paper as:
\begin{align*}
\rmd S_{t} & = (r-\delta_{t})S_{t}\rmd t+\sigma_{a_t} S_{t}\rmd W_{t}^{1} \eqsp,\\
\rmd\delta_{t} & = \kappa (\alpha_{a_t} -\delta_{t})\rmd t+\eta_{a_t} \rmd W_{t}^{2} \eqsp, \quad \rmd\langle W_t^1, W_t^2\rangle = \rho_{a_t} \rmd t\eqsp,
\end{align*}
where $(a_t)_{t\ge 0}$ is a finite state space Markov process.

@@

Z_i = d_{a_{i-1}} + T Z_{i-1} + H_{a_{i-1}} \varepsilon_i\eqsp,

@@

where (with $\barH_{a_{i-1}} \eqdef H_{a_{i-1}}H'_{a_{i-1}}$ and $\Delta = t_i-t_{i-1}$):

@@

\begin{pmatrix} \left[\mu- \alpha_{a_{i-1}} - \sigma^2_{a_{i-1}}/2 \right] \Delta + \alpha_{a_{i-1}}[1-\rme^{-\kappa \Delta}]/\kappa\\
\alpha_{a_{i-1}} [1-\rme^{-\kappa \Delta}] \end{pmatrix}\eqsp,\\
T &\eqdef
\begin{pmatrix} 1 & -[1-\rme^{-\kappa \Delta}]/\kappa \\ 0 & \rme^{-\kappa \Delta}  \end{pmatrix} \eqsp,\\
\barH_{a_{i-1}}(1,1) &= \sigma^2_{a_{i-1}} \Delta + \eta^2_{a_{i-1}}\left\{\Delta  + (1-\rme^{-2\kappa \Delta})/(2\kappa) - 2(1-\rme^{-\kappa \Delta})/\kappa\right\}/\kappa^2 \\
&\hspace{4.4cm} - 2\rho_{a_{i-1}} \eta_{a_{i-1}} \sigma_{a_{i-1}}\left\{ \Delta t_i - (1-\rme^{-\kappa \Delta})/\kappa \right\}/\kappa\eqsp, \\
\barH_{a_{i-1}}(1,2) & = \left(\rho_{a_{i-1}} \eta_{a_{i-1}} \sigma_{a_{i-1}}-\eta^2_{a_{i-1}}/\kappa \right)\left(1-\rme^{-\kappa \Delta}} \right)/\kappa} \\
&\hspace{7.8cm}+ \eta^2_{a_{i-1}}\left(1-\rme^{-2\kappa \Delta}} \right)/(2\kappa^2)\eqsp,\\

@@

\barH_{a_{i-1}}(2,2) &= \eta^2_{a_{i-1}}\left( 1-\rme^{-2\kappa \Delta} \right)/(2\kappa)\eqsp.

@@

A closed form solution for $F_{t_i,m}$  may be written:
